\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-07-14 paper.qmd
still working on hex grid

```{r}
#| label: load-libraries
#| warning: false
#| echo: false
library(quollr)
library(tibble)
library(dplyr)
# remotes::install_github("jlmelville/snedata")
library(snedata)
library(ggflowchart)
library(purrr) ## map function
library(gridExtra) ## for grid.arrange
library(rsample)
library(DT)
library(ggbeeswarm)
library(ggplot2)
library(readr)
library(tidyr)

@@

#### Scaling the data

Because we are working with distances between points, starting with data having a standard scale, e.g. [0, 1], is recommended. When the aspect ratio of the NLDR $(r_1, r_2, ..., r_k)$ is meaningful, the data can be scaled to take this into account. When $k=2$, as in hexagon binning, the default range is $[0, y_{i,\text{max}}], i=1,2$, where $y_{1,\text{max}}=1$ and $y_{2,\text{max}} = \frac{r_2}{r_1}$ (@fig-NLDR-scurve). <!-- \times \frac{2}{\sqrt{3}}$. (The $\frac{2}{\sqrt{3}}$ accounts for the different height ($a_1$) and width ($a_2$) of a regular hexagon.) The scaling of data should take the size of the hexagons into account, but choice of number of bins should. -->

@@

The \gD{} hexagon grid is defined by the number of bins in each direction $(b_1, b_2)$, giving total number of bins as $b = b_1 \times b_2$, and a unique hexagon id, $h = 1, \dots, b$. Each hexagon, $H_h$ is uniquely described by centroid, $C_{h}^{(2)} = (c_{h1}, c_{h2})$.  (Note that, we detail the algorithm for constructing the grid and binning the data because we found that there was no readily available implementation that could produce what was needed for this project.)

To make the grid, a buffer parameter ($q$) is set as a proportion of the plot space. By default this is 0.1. A value should be chosen to provide at buffer of a full hexagon width ($a_1$) and height ($a_2$) for the grid to extend beyond the data. The lower left position where the grid starts is defined as $(s_1, s_2)$, corresponding to the centroid of the lowest left hexagon, which should be below the minimum data value. This is computed as one buffer unit, $q$ below the minimum data value, by default. The user provides the preferred number of bins in the horizontal direction, $b_1$, and the algorithm computes the appropriate number of bins in the vertical direction, $b_2$.

>>>>

```{r}
#| label: load-libraries
#| warning: false
#| echo: false
# remotes::install_github("JayaniLakshika/quollr")
library(quollr)
library(tibble)
library(dplyr)
# remotes::install_github("jlmelville/snedata")
library(snedata)
library(ggflowchart)
library(purrr) ## map function
library(gridExtra) ## for grid.arrange
library(rsample)
library(DT)
library(ggbeeswarm)
library(ggplot2)
library(readr)
library(tidyr)

@@

### Algorithm to represent the model in 2D

#### Scale the data

Because we are working with distances between points, starting with data having a standard scale, e.g. [0, 1], is recommended. The default should take the aspect ratio produced by the NLDR $(r_1, r_2, ..., r_k)$ into account. When $k=2$, as in hexagon binning, the default range is $[0, y_{i,\text{max}}], i=1,2$, where $y_{1,\text{max}}=1$ and $y_{2,\text{max}} = \frac{r_2}{r_1}$ (@fig-NLDR-scurve). If the NLDR aspect ratio is ignored then set $y_ {2,\text{max}} = 1$. <!-- \times \frac{2}{\sqrt{3}}$. (The $\frac{2}{\sqrt{3}}$ accounts for the different height ($a_1$) and width ($a_2$) of a regular hexagon.) The scaling of data should take the size of the hexagons into account, but choice of number of bins should. -->

@@

The \gD{} hexagon grid is defined by the number of bins in each direction $(b_1, b_2)$, giving total number of bins as $b = b_1 \times b_2$, and a unique hexagon id, $h = 1, \dots, b$. Each hexagon, $H_h$ is uniquely described by centroid, $C_{h}^{(2)} = (c_{h1}, c_{h2})$.  (Note that, we explain the algorithm for constructing the grid and binning the data here because, surprisingly, there is no readily available implementation that could produce everything needed for this project.)

To make the grid, a buffer parameter ($q$) is set as a proportion of the plot space. By default this is 0.1. A value should be chosen to provide at buffer of a full hexagon width ($a_1$) and height ($a_2$) for the grid to extend beyond the data, in all directions. The lower left position where the grid starts is defined as $(s_1, s_2)$, and corresponds to the centroid of the lowest left hexagon. This must be smaller than the minimum data value. This is computed as one buffer unit, $q$ below the minimum data value, by default. The user provides the preferred number of bins in the horizontal direction, $b_1$, and the algorithm computes the appropriate number of bins in the vertical direction, $b_2$.

### 2024-07-14 paper.tex
still working on hex grid

\subsubsection{Scaling the data}\label{scaling-the-data}

Because we are working with distances between points, starting with data
having a standard scale, e.g.~{[}0, 1{]}, is recommended. When the
aspect ratio of the NLDR \((r_1, r_2, ..., r_k)\) is meaningful, the
data can be scaled to take this into account. When \(k=2\), as in
hexagon binning, the default range is \([0, y_{i,\text{max}}], i=1,2\),
where \(y_{1,\text{max}}=1\) and \(y_{2,\text{max}} = \frac{r_2}{r_1}\)
(Figure~\ref{fig-NLDR-scurve}).

@@

The \gD{} hexagon grid is defined by the number of bins in each
direction \((b_1, b_2)\), giving total number of bins as
\(b = b_1 \times b_2\), and a unique hexagon id, \(h = 1, \dots, b\).
Each hexagon, \(H_h\) is uniquely described by centroid,
\(C_{h}^{(2)} = (c_{h1}, c_{h2})\). (Note that, we detail the algorithm
for constructing the grid and binning the data because we found that
there was no readily available implementation that could produce what
was needed for this project.)

To make the grid, a buffer parameter (\(q\)) is set as a proportion of
the plot space. By default this is 0.1. A value should be chosen to
provide at buffer of a full hexagon width (\(a_1\)) and height (\(a_2\))
for the grid to extend beyond the data. The lower left position where
the grid starts is defined as \((s_1, s_2)\), corresponding to the
centroid of the lowest left hexagon, which should be below the minimum
data value. This is computed as one buffer unit, \(q\) below the minimum
data value, by default. The user provides the preferred number of bins
in the horizontal direction, \(b_1\), and the algorithm computes the
appropriate number of bins in the vertical direction, \(b_2\).

>>>>

\subsection{Algorithm to represent the model in
2D}\label{algorithm-to-represent-the-model-in-2d}

\subsubsection{Scale the data}\label{scale-the-data}

Because we are working with distances between points, starting with data
having a standard scale, e.g.~{[}0, 1{]}, is recommended. The default
should take the aspect ratio produced by the NLDR
\((r_1, r_2, ..., r_k)\) into account. When \(k=2\), as in hexagon
binning, the default range is \([0, y_{i,\text{max}}], i=1,2\), where
\(y_{1,\text{max}}=1\) and \(y_{2,\text{max}} = \frac{r_2}{r_1}\)
(Figure~\ref{fig-NLDR-scurve}). If the NLDR aspect ratio is ignored then
set \(y_ {2,\text{max}} = 1\).

@@

The \gD{} hexagon grid is defined by the number of bins in each
direction \((b_1, b_2)\), giving total number of bins as
\(b = b_1 \times b_2\), and a unique hexagon id, \(h = 1, \dots, b\).
Each hexagon, \(H_h\) is uniquely described by centroid,
\(C_{h}^{(2)} = (c_{h1}, c_{h2})\). (Note that, we explain the algorithm
for constructing the grid and binning the data here because,
surprisingly, there is no readily available implementation that could
produce everything needed for this project.)

To make the grid, a buffer parameter (\(q\)) is set as a proportion of
the plot space. By default this is 0.1. A value should be chosen to
provide at buffer of a full hexagon width (\(a_1\)) and height (\(a_2\))
for the grid to extend beyond the data, in all directions. The lower
left position where the grid starts is defined as \((s_1, s_2)\), and
corresponds to the centroid of the lowest left hexagon. This must be
smaller than the minimum data value. This is computed as one buffer
unit, \(q\) below the minimum data value, by default. The user provides
the preferred number of bins in the horizontal direction, \(b_1\), and
the algorithm computes the appropriate number of bins in the vertical
direction, \(b_2\).
